---
title: "Problem Sets 1"
output:
  word_document: default
  html_document: default
date: "2024-02-17"
resources: "https://stackoverflow.com/questions/6286313/remove-an-entire-column-from-a-data-frame-in-r"
---

# The first Markdown report for problems set one
## By : Decker Mecham




# Question 1

# A.  Check current working directory
```{r}
getwd()
```


# B.  Read the dataset into R and call it "movies"
```{r}
movies <- read.csv("datasets/IMDB_movies.csv")
```
# C . Dimensions of the dataset
```{r}
dim(movies)
```

# D. Names of the variables in the dataset
```{r}
names(movies)
```
```{r}
library(tidyverse)


movies %>% glimpse()
```
# E . char variables below
movie_title, director_name, country, language, content_rating, genres, color, plot_keywords, actor_1_name, actor_2_name

# F. Show the first 5 rows of the dataset
```{r}
slice(movies, 1:5)
```
# G. Arrange the dataset in descending order by gross (total box office revenue in $USD) and show the ten movies with the largest box office revenue
```{r}
top_10_grossing = arrange(movies, (-gross)) %>% slice(1:10)
top_10_grossing
top_10_grossing$director_name[1]
```




# H. Remove the variable facenumber_in_poster and store the smaller dataset as movies_sub, and check dimensions to ensure movies_sub has one fewer variable
```{r}
movies_sub <- select(movies, -facenumber_in_poster)
dim(movies_sub)
dim(movies)
```
# I. Number of movies rated as PG-13, and Average revenue (gross) for PG-13 movies
```{r}
pg_13_movies = filter(movies, content_rating == "PG-13")
nrow(pg_13_movies)
mean(pg_13_movies$gross, na.rm = TRUE)
```


# Question 2 (first load library for ggplot2)
```{r}
library(ggplot2)
```
# A. Create scatter plot of IMDB score vs. gross
```{r, echo = FALSE}
ggplot(movies, aes(x = imdb_score, y = gross)) +
  geom_point() +
  labs(x = "IMDB Score", y = "Gross (USD)") +
  ggtitle("IMDB Score vs. Gross") +
  theme_minimal()
```

# B. add alpha 1/10
```{r,echo = FALSE}
ggplot(movies, aes(x = imdb_score, y = gross)) +
  geom_point(alpha = 1/10) +
  labs(x = "IMDB Score", y = "Gross (USD)") +
  ggtitle("IMDB Score vs. Gross") +
  theme_minimal()
```
 
 
# C. Create a new variable for gross in millions of dollars 
```{r}
movies <- movies %>% mutate(gross_M = gross / 1000000)
```

# D. Create scatter plot of IMDB score vs. gross in millions with smoothing line
```{r, echo = FALSE}
ggplot(movies, aes(x = imdb_score, y = gross_M)) +
  geom_point(alpha = 1/10) +
  geom_smooth(method = "lm", color = "blue") +
  labs(x = "IMDB Score", y = "Gross (Millions of USD)") +
  ggtitle("IMDB Score vs. Gross (in Millions)") +
  theme_minimal()
```

# E. Filter movies by James Cameron and create scatter plot of IMDB score vs. gross in millions for James Cameron films with movie titles as points
```{r}
cameron_movies <- filter(movies, director_name == "James Cameron")
```

```{r, echo = FALSE}
ggplot(cameron_movies, aes(x = imdb_score, y = gross_M)) +
  geom_point(color = "blue") +
  geom_text(aes(label = movie_title), color = "blue") +
  ggtitle("James Cameron Films: IMDB Score vs. Gross (in Millions)") +
  theme_minimal()
```


# F. more reasonable x and y axis with xlim 
```{r, echo = FALSE}
ggplot(cameron_movies, aes(x = imdb_score, y = gross_M)) +
  geom_point(color = "blue") +
  geom_text(aes(label = movie_title), color = "blue") +
  xlim(7, 9) + ylim(0, 800) +
  ggtitle("James Cameron Films: IMDB Score vs. Gross (in Millions)") +
  theme_minimal()
```


# G.  better name for x and y axis with xlab and ylab or using labs is possible as well. 
```{r, echo = FALSE}
ggplot(cameron_movies, aes(x = imdb_score, y = gross_M)) +
  geom_point(color = "blue") +
  geom_text(aes(label = movie_title), color = "blue") +
  xlim(7, 9) + ylim(0, max(cameron_movies$gross_M)) +
  xlab("IMDB Score") + ylab("Gross (Millions of USD)") +
  ggtitle("James Cameron Films: IMDB Score vs. Gross (in Millions)") +
  theme_minimal()
```



